---
title: "Initial Korean covid 19"
author: "Omar Challan"
date: "03/19/2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

## Methods

Based on the problem description, it appears that the problem can be solved as a categorical – categorical problem as we are comparing two categorical data types. The first is sex, and the second is the state of the patient. 

## Graphical results - use ggplot commands to do this  


```{r}
dd2 <- kRdata %>% group_by(result,sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

From the previously shown two graphs, it shows that more women are getting Covid than men. However, more men die from Covid than women. 

  
## Numerical results

```{r}
table1 <- xtabs(~result + sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The numerical data confirms the results we concluded from the previous graphical data, which was that more women get Covid than men. However, the percentage of men who died from Covid is higher than women. 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

After running these inferential tests, we found that the P value is 0.0009, which is far lower than 0.05 and based on these results, we can say that males are more than twice as likely to die from Covid than females.

## Conclusions

Looking at this problem, we have determined that it's a categorical – categorical problem, as we are comparing two categorical Data types which are sex (Male – female) and the status of the patient which is either alive or dead. After showing the graphical data that compares males and females contracting and dying from Covid. We concluded that more females got Covid. However, more males died from Covid. This conclusion was further confirmed by the numerical data results and the inferential tests, as the P value was far lower than the standard 0.05. Therefore, we can conclude that the death rates of 
patients is not independent from there sex. 

